
--- 
title: "Reproducible Research - Course Project 1" 
output: 
  html_document: 
    keep_md: true 
---
### Loading and Preprocessing the Data

```{r echo=TRUE}
activitydata <- read.csv("activity.csv")
colnames(activitydata) <- c("steps","date","interval")
subset_activitydata <- activitydata[which(activitydata$steps != "NA"), names(activitydata) %in%
                                            c("steps","date","interval")]
head(subset_activitydata)
```


### What is mean total number of steps taken per day?

```{r echo=TRUE}
library(ggplot2)
steps_by_day <- aggregate(subset_activitydata$steps, by=subset_activitydata["date"], FUN=sum)
head(steps_by_day)
ggplot(steps_by_day, aes(x=x)) + geom_histogram()
mean(steps_by_day$x)
median(steps_by_day$x)
```

### What is the average daily activity pattern?
```{r echo=TRUE}
library(ggplot2)
average_steps_per_interval <- aggregate(subset_activitydata$steps, by=subset_activitydata["interval"], FUN=mean)
head(average_steps_per_interval)
ggplot(average_steps_per_interval,
              aes(x = interval, y = x)) + geom_line()
average_steps_per_interval[which.max(average_steps_per_interval$x),]
```

### Inputing Missing Values
### Rows with NA
```{r echo=TRUE}
count_all_rows <- nrow(activitydata)
count_filled_rows <- nrow(subset_activitydata)
return(count_all_rows-count_filled_rows)
```

### Fill NAs with Average Value for Interval
```{r echo=TRUE}
merge1 <- merge(activitydata, average_steps_per_interval, by="interval")
for (i in 1:nrow(merge1)){
        if (is.na(merge1[i,2])){ 
                merge1[i,2] <- merge1[i,4]
        }
        else {
                merge1[i,2] <- merge1[i,2]
                }
}
head(merge1)
steps_by_day_2 <- aggregate(merge1$steps, by=merge1["date"], FUN=sum)
head(steps_by_day_2)
ggplot(steps_by_day_2, aes(x=x)) + geom_histogram()
mean(steps_by_day_2$x)
median(steps_by_day_2$x)
```
##### Yes, the values differ from the estimates from the first part of the assignment. The impact is that now the median number of steps and mean number of steps are the same.

### Are there differences in activity patterns between weekdays and weekends?
```{r echo=TRUE}
for (i in 1:nrow(merge1)){
        merge1[i,5] <- weekdays(as.Date(merge1[i,3]), abbreviate = FALSE)}
steps_by_day_3 <- aggregate(merge1$steps, by=merge1["V5"], FUN=mean)
print(steps_by_day_3)
merge2 <- merge(merge1, steps_by_day_3, by="V5")
subset_merge2_weekend <- merge2[which((merge2$V5 == "Saturday") | (merge2$V5 == "Sunday")), names(merge2) %in%
                                            c("V5","interval","steps","date","x.x","x.y")]
subset_merge2_weekday <- merge2[which((merge2$V5 != "Saturday") | (merge2$V5 != "Sunday")), names(merge2) %in%
                                            c("V5","interval","steps","date","x.x","x.y")]

steps_by_interval <- aggregate(subset_merge2_weekend$steps, by=subset_merge2_weekend["interval"], FUN=mean)
colnames(steps_by_interval) <- c("interval","steps")

steps_by_interval2 <- aggregate(subset_merge2_weekday$steps, by=subset_merge2_weekday["interval"], FUN=mean)
colnames(steps_by_interval2) <- c("interval","steps")

plot1 <- ggplot(steps_by_interval,
              aes(x = interval, y = steps)) + geom_line() + ggtitle("Weekend")

plot2 <- ggplot(steps_by_interval2,
              aes(x = interval, y = steps)) + geom_line() + ggtitle("Weekday")

library(gtable)
plot1 <- ggplotGrob(plot1)
plot2 <- ggplotGrob(plot2)
g <- rbind(plot1, plot2, size = "first")
grid::grid.draw(g)
```
